Performance of Individual Audit Filters in Predicting Opportunities for Improvement

### 2024-06-08 manuscript.Rmd
update

Discussion
==========
This study showed that AFs do not perform well in predicting OFIs. No single AF presented with a performance of value in predicting OFI in trauma care patients. Some AF showed a correlation with OFI but not enough to be considered useful.

Patient characteristics are similar to previous studier showing twofold prevalence of male trauma patients compared to females [@roth2018global;@SweTrau2022Arsrapport]. Previous studies have shown

A previous study conducted at Karolinska University Hospital by Albaaj
et al. [@albaaj2023patient] sought to identify patient factors
associated with OFIs and found that a higher ISS, longer times to CT,
longer times to first intervention and higher levels of care was
correlated with OFIs. These findings correspond with our analysis of
patient characteristics and AFs linked to these factors.

Higher ISS and higher levels of care in patients with OFIs as opposed to
patients without OFIs could be explained by the greater complexity of
such trauma cases. Greater complexity requires more interventions which
could allow for more mistakes [@albaaj2023patient].

@@

AFs generally exhibited a high specificity relative to their sensitivity
which may be attributed to the outcome being rather uncommon, whilst
every individual AF misses the majority of OFIs. This could suggest that
the AFs are too rigid in their criteria. On the contrary, a too
inclusive criteria might flag several patients, who do not necessarily
have an OFI [@willis2008interpreting].

In 2008, Willis et al. [@willis2008interpreting] assessed 14 audit
filters, finding only three to be associated with an increased risk of
poor outcomes. One AF, "GCS\<9 and not intubated", also used in this
study, similarly showed no significant ability to predict OFIs. An
explanation for this could be that an abnormal GCS as a result from
intoxication is not comparable to an abnormal GCS due to head injury for
example, suggesting that the AF might be more reflective of clinical
decision rather than quality of care [@willis2008interpreting]. However,
according to Sanddal et al. [@sanddal2011analysis] airway management is
associated with OFI suggesting that an alternative audit filter related
to airway management might potentially perform differently.

Both AFs measuring delay: "\>60 until first intervention" and "\>30min
until first CT", showed a correlation with OFI. This is comparable to
what Teixeira et al. [@teixeira2007preventable] demonstrated, with delay
correlating to preventable death. Furthermore, Ghorbani et al.
[@ghorbani2018analysis] identified delay in treatment, delay to CT, and
clinical judgement errors as common errors in trauma care. A more recent
study in Japan by Yamamoto et. al [@yamamoto2023immediate] showed a
significant reduction in mortality when patients received a whole-body
CT within 10 minutes of arrival at the ED. It is important to note that
OFI in this study was not only a reflection of mortality but a broader
concept encompassing different aspects of care quality. If anything, it
highlights the difference between hospitals, which AFs are in use, and
how they are being assessed.

AFs that present with a high risk of inevitable death showed poor"Dead
at 30 days" and "CPR and thoracotomy", could be explained by inevitable
mortality, thus presenting with a smaller time frame for mistakes to be
made.

@@

AFs are static in their criteria, contrary to health care in general.
Consequently, the utility of AFs may diminish if confronted with
emerging challenges or upon resolution of the quality gaps they were
designed for. The static nature of AFs highlight their need for
consistent and evidence-based review to determine current usefulness
[@santana2014development]. However, it is worth pointing out that the
review process might present with costs greater than the potential
benefit from AFs.

AFs face noteworthy challenges due to over-reliance, bias and
confounders. Firstly, while AFs may help in determining the existence of
OFI, they could overshadow other factors that may contribute to that
analysis, leading to over-reliance. Secondly, devising AFs comes with an
unintended bias. They might disproportionally affect certain groups of
people failing to recognise patient cases that fall outside the scope of
AF criteria [@willis2008interpreting]. For example, AFs primarily focus
on aspects of trauma care mostly relevant to medical doctors despite the
interdisciplinary nature of trauma care. This is especially
counterintuitive given the broad definition of OFIs. Lastly, confounding
variables adds to the uncertainty of AFs performance. For example,
patients receiving the same interventions at different time points may
do so as a function of injury severity not quality of care
[@willis2008interpreting].

@@

Opportunity for improvement, while dichotomously defined, includes a
diverse set of outcomes. This convolutes the process of creating AFs
that perform well in predicting OFIs. The current screening system for
OFIs might also introduce bias since the AFs would favour the
identification of some, but not all, errors.

Each AF was individually related to OFI. Consequently, even if an AF
demonstrated a true positive correlation with an OFI, instances where
the same patient case had multiple AFs introduce a degree of uncertainty
regarding which specific AF accurately predicted the outcome.

@@

Our findings suggest that AFs perform poorly in predicting OFIs. The AFs
with performance exceeding coincidence (AUC = 0.5) might be a beneficial
addition to an already existing quality improvement programme or could
be used to fine-tune AFs already in use to plug hospitals' quality gaps.
Due to the study being conducted in a single, trauma level I centre in
Stockholm it is unrealistic to assume applicability of these findings at
other hospitals, in other countries, facing unique challenges
[@world2009guidelines]. For example, Low and middle income countries
seem to benefit from certain AFs that high income countries do not and
vice versa [@berg2022perceived].

>>>>

Discussion
==========
This study showed that AFs do not perform well in predicting OFIs. 
AFs generally exhibited a high specificity relative to their sensitivity
which may be attributed to the outcome being rather uncommon, whilst
every individual AF misses the majority of OFIs. This could suggest that
the AFs are too rigid in their criteria. On the contrary, a too
inclusive criteria might flag several patients, who do not necessarily
have an OFI [@willis2008interpreting].

AFs are static in their criteria, contrary to health care in general.
Consequently, the utility of AFs may diminish if confronted with
emerging challenges or upon resolution of the quality gaps they were
designed for. This highlights their need for
consistent and evidence-based review to determine current usefulness
[@santana2014development]. However, such
review process might present with costs greater than the potential
benefit from AFs.

While AFs may help determining
OFIs, they could overshadow other factors contributing to that
analysis, resulting in over-reliance. AFs might disproportionally affect certain groups of
people failing to recognise patient cases that fall outside the scope of
AF criteria [@willis2008interpreting]. For example, AFs primarily focus
on aspects of trauma care mostly relevant to medical doctors despite the
interdisciplinary nature of trauma care. This is especially
counterintuitive given the broad definition of OFIs. Lastly, confounding
variables adds to the uncertainty of AFs performance. For example,
patients receiving the same interventions at different time points may
do so as a function of injury severity not quality of care
[@willis2008interpreting].

Higher ISS, longer times to CT, longer times to first intervention and higher levels of care were correlated with OFIs, corresponding with a previous study conducted at Karolinska University Hospital by Albaaj
et al. [@albaaj2023patient]. Greater severity and higher levels of care call for greater complexity, requiring more interventions, allowing for more mistakes. This could explain their correlation with OFI[@albaaj2023patient].

@@

The AF "GCS<9 and not intubated" showed no significant ability to predict OFIs, similarly shown by Willis et al. [@willis2008interpreting] in 2008. An
explanation for this could be that an abnormal GCS as a result from
intoxication versus head injury for example are not comparable, suggesting that the AF might be more reflective of clinical
decision rather than quality of care [@willis2008interpreting]. However, according to Sanddal et al. [@sanddal2011analysis] airway management was
associated with OFI, suggesting that an alternative AF related
to airway management might perform differently.

Both AFs measuring delay: ">60 until first intervention" and ">30min
until first CT", showed a correlation with OFI. This is comparable to
what Teixeira et al. [@teixeira2007preventable] demonstrated, with delay
correlating to preventable death. Furthermore, Ghorbani et al.
[@ghorbani2018analysis] identified delay in treatment and delay to CT as common errors in trauma care. A more recent
study in Japan by Yamamoto et. al [@yamamoto2023immediate] showed a
significant reduction in mortality when patients received a whole-body
CT within 10 minutes of arrival at the ED. Notably,
OFI in this study was not soley a reflection of mortality but also morbidity. Rather, it
highlights the difference between hospitals, which AFs are in use, and
how they are assessed.

AFs that present with a high risk of inevitable death: "Dead
at 30 days" and "CPR and thoracotomy" showed no correlation with OFI. This could be explained by inevitable
mortality, thus presenting with a smaller time frame for mistakes to be
made.

@@

@@

OFI, while dichotomously defined, includes a
diverse set of outcomes. This convolutes the process of creating AFs
that perform well in predicting OFIs. AFs would favour the
identification of some, but not all, errors.

Each AF was individually related to OFI. Consequently, if an AF
demonstrated a true positive correlation with an OFI, instances where
the same patient case had multiple AFs introduce uncertainty
regarding which specific AF truthfully predicted the OFI.

@@

Our findings suggest that AFs perform poorly in predicting an OFI. The AFs
with performance exceeding coincidence (AUC = 0.5) might be a beneficial
addition to an already existing QI programme or
be used to fine-tune AFs already in use to plug hospitals' quality gaps.
Due to the study being conducted in a single, trauma level I centre in
Stockholm it is unrealistic to assume applicability of these findings at
other hospitals facing unique challenges
[@world2009guidelines]. For example, Low and middle income countries
seem to benefit from certain AFs that high income countries do not and
vice versa [@berg2022perceived].
